---
title: "data_exploration"
format: html
editor: visual
---

## Initial Data Exploration

To start, we will do the following in a 'static' manner: - Produce numerical and graphical summaries to investigate the data. These should include: – One-way and Two-way contingency tables

```         
– Numerical summaries (means, medians, sds, etc.) for quantitative variables at levels of categorical variables
– Create at least six plots.
  - At least four of these should display multivariate information via the type of graph, by utilizing coloring, grouping, etc.
  - All plots should have nice labels and titles.
  - Some kind of faceting should be used somewhere.
  - At least one plot should be a plot that we didn’t cover in class (say a heatmap or something like that (depends on your data)
    lots of good examples to consider here https://exts.ggplot2.tidyverse.org/gallery/)
```

First, we will read in our data.

```{r}
#| warning: FALSE
library(tidyverse)
library(dplyr)
library(readxl)
```

```{r}
superstore_data <- read_excel("US_Superstore_data.xls", sheet = "Orders")
head(superstore_data)
```

### Numerical Summaries

Now, let's look at some numerical summaries to investigate our US Superstore data. First, let's investigate the different sections from where orders are placed. The variable we will use is called `Segment` and the orders can be placed from the following: `Consumer`, `Corporate`, or `Home Office`. By observing the one-way contingency table below, we can see that a majority of orders are placed from a consumer.

```{r One-way contingency table}
table(superstore_data$Segment)
```

Let's add in an extra factor to this piece of data. Across all sections from where orders are placed, we can see that items in the `Office Supplies` category are the most common order placed.

```{r Two-way contingency table}
table(superstore_data$Segment, superstore_data$Category)
```

Let's also find some other numerical summaries (means, medians, sds, etc.) for quantitative variables at levels of categorical variables. Below, we will see some summaries for `Sales` (the selling price of each product) and `Profit` (the profit gained on each product), as they relate to each order **category**. We can see that the `Technology` category has the highest priced products on average, as well as the highest amount of profit for products sold.

```{r Numerical summaries}
superstore_data |>
  group_by(Category) |>
  summarise(
    mean_sales = mean(Sales, na.rm = TRUE),
    median_sales = median(Sales, na.rm = TRUE),
    sd_sales = sd(Sales, na.rm = TRUE),
    
    mean_profit = mean(Profit, na.rm = TRUE),
    median_profit = median(Profit, na.rm = TRUE),
    sd_profit = sd(Profit, na.rm = TRUE)
  )
```

### Graphical Summaries

Next, let's create some graphical summaries and plots to help visualize our data. We will cover multivariate information, facet by some categorical variable, and end with a unique plot!
